Feature Selection and Classification of \\ Spam on Social Networking Sites

### 2012-05-09 docs/paper/paper.tex
update

A number of researchers have focused on collecting, identifying features 
and classifying various genres of spam on social networks. Zinman and 
Donath \cite{zinman} extract bundles of profile-based and comment-based 
features from MySpace profiles, but the relatively poor performance of their classifier highlights 
the difficulties in manual classification social network spam. Several 
studies take the approach of baiting spammers with social ``honeypots", profiles 
created with the sole intent of attracting spam.\cite{stringhini, lee} They 
then use the data collected to train classifiers with features including 
friend request rate and ratios of URLs to text. Webb et al. \cite{webb} use 
the honeypot approach as well and provide examples of various types of spammers, 
the typical demographics of their profiles as well as the web pages that they tend to advertise.

Gao et al. \cite{gao} look at Facebook wall posts, analyzing temporal properties, 
URL characteristics, post ratios and other features of malicious accounts. They 
also pinpoint various spam ``campaigns" based on products advertised in a given 
time frame. They note that spam on Facebook often exhibits burstiness and is mainly 
sent from compromised accounts. Benevenuto et al. \cite{benevenuto} identify 
social attributes of spam and ham on video SNSs (in this case, YouTube), including video 
view counts, comment counts and user public profile attributes. They then use a 
support vector machine (SVM) for classification, yielding 96\% accuracy in detecting 
advertisers (``promoters"), but only accurately identifying 57\% of examples of more general spam.

@@

De Wang et al. \cite{wang} propose a cross-site spam detection framework to share 
spam data across all social networking sites, building classifiers to identify spam 
in profiles, messages and web pages. This multi-pronged approach lends itself to 
associative classification, in which, for example, a message would be classified as 
spam if it contained a link to a web page that had a high probability of being spam. 
Unfortunately, the differing characteristics of various social networks \eg the length 
of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora 
across diverse sites.

Facebook \cite{stein} provides an overview of their ``immune system" defenses 
against phishing, fraud and spam. The system is composed of classifier services, an 
ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare 
features for classification and a policy engine to take action on suspected misuse. While 
the discussion remains high-level and includes few implementation particulars, it does 
include significant detail on the various types and characteristics of undesirable 
activity on the site, including fake profiles, harassment, compromised accounts, malware 
and spam.

@@

Much work has been done on protecting traditional email systems from spam. 
Blanzieri \cite{blanzieri} offers a comprehensive overview of machine learning techniques 
that can be applied to email filtering. Hao et al. \cite{hao} describe a reputation 
engine based on lightweight features such as geographic distance between sender and 
receiver, geolocation anomalies and diurnal patterns. While the target was conventional 
spam, these and similar features are applicable to spam on SNSs as well.

@@

Blog comment spam have also attracted considerable attention from researchers who 
have applied machine learning \cite{kolari, nag} and NLP \cite{mishne} techniques to the 
problem. Likewise, Markines et al. \cite{markines} apply similar techniques to spam on 
social bookmarking sites.

@@

Many of the data mining algorithms used to detect spam and patterns of misuse 
on SNSs are designed with the assumption that the data and the classifier are 
independent. However, in the case of spam, fraud and other malicious content, 
users will often modify their behavior to evade detection, leading to degraded 
classifier performance and the need to re-train classifiers frequently. Several 
researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a 
modified Naive Bayes classifier to detect and reclassify data taking into account 
the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} 
provide a framework for reverse engineering a classifier to determine whether 
an adversary can efficiently learn enough about a classifier to effectively defeat it.

@@

For this project, we had unlimited access to data from InterPals, a SNS for users who 
wish to communicate with others from other countries, whether for language practice, cultural 
exchange or friendship. Users sign up by completing a registration form with 
information about themselves, including age, sex and location. After registering, users 
can expand their profile page to include descriptions, interests, languages they speak, etc. 
and upload photographs. After clicking on a link sent to their email address, a user can 
begin to interact with others on the site via private message, public ``wall" posts, comments 
on photos and on a bulletin board system.

@@

Currently, the site combats spam and other Terms of Service violations 
through volunteer moderators. Users can report content, including private messages, 
profiles and photographs to moderators, using a form that includes a drop-down menu of 
pre-selected reasons with the option to add a more detailed message in a text field. 
Likewise, the message interface allows a user to report a private message spam with a single 
click. Moderators have access to a queue of these reports. In addition to the material 
being reported, moderators are able to make decisions based on the data from the reported 
user's account, including outgoing private messages, IP addresses, as well as a list of 
other users who have logged in from the same computer (determined via IP address as well 
as by browser fingerprints). Moderators can then decide to delete the user, send a warning, 
or clear reports on a user. All moderator actions require them to annotate their decision with a 
brief log message. When deleting a user, moderators have the option of flagging the 
reported message (if there is one) as spam, or of flagging all of the user's outgoing private messages 
as spam.

Other anti-spam measures include widespread use of CAPTCHAs across the site and 
frequency caps. Short-window frequency caps are in place for all users, limiting the 
number of messages that can be sent per short time interval (1, 5 and 10 minutes). New users 
are also subject to a per-day cap on the number of unique users with whom they are able to initiate 
contact.

@@

At the beginning of this research project, slightly two million messages had 
been flagged by moderators as spam. We extracted the contents of exactly 2 million 
spam-labeled messages from accounts deleted by moderators between October 2011 
and March 2012. As moderators can flag a user's entire list of outgoing messages 
as spam on deletion, the earliest sent dates of some messages in our data set begin in 
May 2010.

InterPals stores private messages and user account data in a number of separate MySQL tables. 
Account information for deleted users is stored for six months, so account details for all spam 
and ham accounts was available. Extracting the data of interest to this project required 
dumping the contents from a query with multiple joins and aggregations across 8 tables with a 
combined size of 154 GB. To minimize the impact on the load of the site's production database 
instances, we took an LVM snapshot of slave instance and cloned the database to a spare server. 
In addition to using SQL to extract the data, we wrote Perl scripts to clean the data, tokenize 
messages and prepare data for further processing. The methodology section (X) offers more details 
on precisely which items of data were collected and why.

@@

Users are free to choose any country on their profile. We record each IP address from which a 
user logs in to the site. By examining the distribution of countries (using the MaxMind GeoIP 
database) associated with the unique IP addresses (Figure \ref{fig:cip}), we see that the top 
countries for spammers are significantly different, while the top ham countries are 
virtually unchanged. Furthermore, only 30\% of the IP addresses of spam users who 
claimed to be from the United States actually mapped back to a US-based ISP, with a combined 
46\% indicating a Ghana, Nigeria and Senegal ISP (Figure \ref{fig:usaspam}). This contrasts 
with the 73\% that US-based IP addresses comprised for ham users whose profiles stated that 
they were in the United States (Figure \ref{fig:usaham}). During feature extraction, we also 
generated a boolean feature (IP mismatch) indicating whether the IP-detected country and 
the profile country matched.

@@

It is also anecdotally clear from moderator log messages that spammers re-use their profile photos and 
these photos are often of celebrities or simply stock photos. Fingerprinting and perceptual hash techniques 
such as Marr wavelet, discrete cosine transform (DCT), color histograms, etc., which provided by the 
pHash library, \footnote{http://www.phash.org} seem to promising in detecting image re-use.

>>>>

A number of researchers have focused on collecting, identifying features 
and classifying various genres of spam on social networks. Zinman and 
Donath \cite{zinman} extract bundles of profile-based and comment-based 
features from MySpace profiles, but the relatively poor performance of their classifier highlights 
the difficulties in manual classification of social network spam. Several 
studies \cite{stringhini, lee} take the approach of baiting spammers with social ``honeypots", profiles 
created with the sole intent of attracting spam. They 
then use the data collected to train classifiers with features including 
friend request rate and ratios of URLs to text. Webb et al. \cite{webb} use 
the honeypot approach as well and provide examples of various types of spammers, 
the typical demographics of their profiles, as well as the web pages that they tend to advertise.

Gao et al. \cite{gao} look at Facebook wall posts, analyzing temporal properties, 
URL characteristics, post ratios and other features of malicious accounts. They 
also pinpoint various spam ``campaigns" based on products advertised in a given 
time frame. They note that spam on Facebook often exhibits burstiness and is mainly 
sent from compromised accounts. Benevenuto et al. \cite{benevenuto} identify 
social attributes of spam and ham on video SNSs (in this case, they scraped data from 
YouTube), including video view counts, comment counts and user public profile attributes. 
They then use a support vector machine (SVM) for classification, yielding 96\% accuracy in detecting 
advertisers (``promoters"), but only accurately identifying 57\% of examples of more general spam.

@@

De Wang et al. \cite{wang} propose a cross-site spam detection framework to share 
spam data across all social networking sites, building classifiers to identify spam 
in profiles, messages and web pages. This multi-pronged approach lends itself to 
associative classification, in which, for example, a message would be classified as 
spam if it contains a link to a web page that has a high probability of being spam. 
Unfortunately, the differing characteristics of various social networks \eg the length 
of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora 
across diverse sites.

Facebook \cite{stein} provides an overview of their ``immune system" defenses 
against phishing, fraud and spam. The system is composed of classifier services, an 
ML-derived Feature Extraction Language (FXL), ``feature loops" or services that aggregate 
and prepare features for classification and a policy engine to take action on suspected misuse. While 
the discussion remains high-level and includes few implementation particulars, it does 
include significant detail on the various types and characteristics of undesirable 
activity on the site, including fake profiles, harassment, compromised accounts, malware 
and spam.

@@

Much work has been done on protecting traditional email systems from spam. 
Blanzieri \cite{blanzieri} offers a comprehensive overview of machine learning techniques 
that can be applied to email filtering. Hao et al. \cite{hao} describe a reputation 
engine based on lightweight features such as geographic distance between sender and 
receiver and diurnal patterns. While the target was conventional 
spam, monitoring sender reputation and using similar features (\eg time-of-day when messages 
were sent), seesm applicable to spam on SNSs as well.

@@

Blog comment spam have also attracted considerable attention from researchers who 
have applied machine learning \cite{kolari, nag} techniques including SVMs and Bayesian classifiers.
Mishe et al.\cite{mishne} employ language modeling to find semantic discrepancies between 
the blogs on which link spam comments are posted and the target sites (which might, for example, 
contain adult content). Likewise, Markines et al. \cite{markines} apply similar techniques 
including SVMs and boosting (AdaBoost) to spam on social bookmarking sites.

@@

Many of the data mining algorithms used to detect spam and patterns of misuse 
on SNSs are designed with the assumption that the data and the classifier are 
independent. However, in the case of spam, fraud and other malicious content, 
users will often modify their behavior to evade detection, leading to degraded 
classifier performance and the need to re-train classifiers frequently. Several 
researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a 
modified naive Bayes classifier to detect and reclassify data taking into account 
the optimal modification strategy that an adversary could choose, given full knowledge of 
the classifier model and parameters, but without the ability to tamper with the 
classifier's training data. The authors show that the optimized classifier's counter-strategy 
is substantially more effective than a standard classifers in the cases examined: 
adversaries adding words, adding length and substituting words with synonyms. 
Lowd and Meek \cite{lowd} provide a framework for reverse engineering a 
classifier to determine whether an adversary can efficiently learn enough 
about a classifier to effectively defeat it.

@@

For this project, we had unlimited access to data from InterPals, a site for users who 
wish to communicate with people from other countries, whether for language practice, cultural 
exchange or friendship. Users sign up by completing a registration form with 
information about themselves, including age, sex and location. After registering, users 
can expand their profile page to include self-descriptions, interests, languages they speak,  
photographs, etc. After clicking on an activation link sent to their email address, a user can 
begin to interact with others on the site via private message, public ``wall" posts, comments 
on photos and on a bulletin board system.

@@

Currently, the site combats spam and other Terms of Service violations 
through volunteer moderators. Users can report content, including private messages, 
profiles and photographs, to moderators by using a form that includes a drop-down menu of 
pre-selected reasons with the option to add a more detailed message in a text field. 
Likewise, the message interface allows a user to report a private message as spam with a single 
click. Moderators have access to a queue of these reports. In addition to the material 
being reported, moderators are able to make decisions based on the data from the reported 
user's account, including outgoing private messages, IP addresses, as well as a list of 
other users who have logged in from the same computer (determined via IP address, cookies, as well 
as by rudimentary browser fingerprints). Moderators can then decide to delete the user, send a warning, 
or clear reports on a user. All moderator actions require them to annotate their decision with a 
brief log message. When deleting a user, moderators have the option of flagging the 
reported message (if there is one) as spam or of flagging all of the user's outgoing private messages 
as spam.

Other anti-spam measures include widespread use of CAPTCHAs across the site and 
frequency caps on activities that involve contact with other users. Short-window 
frequency caps are in place for all users, limiting the 
number of messages that can be sent per short time interval (1, 5 and 10 minutes). New users 
are also subject to a per-day cap on the number of unique users with whom they are able to initiate 
contact.

@@

At the beginning of this research project, slightly over two million messages had 
been flagged by moderators as spam. We extracted the contents of exactly 2 million 
spam-labeled messages from accounts deleted by moderators between October 2011 
and March 2012. As moderators can flag a user's entire list of outgoing messages 
as spam on deletion, the earliest sent dates of some messages in our data set begin in 
May 2010.

InterPals stores private messages and user account data in a number of separate MySQL tables. 
Account information for deleted users is stored for six months, so account details for all spam 
and ham accounts were available. Extracting the data of interest to this project required 
dumping the contents from a query with multiple joins and aggregations across 8 tables with a 
combined size of 154 GB. To minimize the impact on the load of the site's production database 
instances, we took an LVM snapshot of slave instance and cloned the database to a spare server. 
In addition to using SQL to extract the data, we wrote Perl scripts to clean the data, tokenize 
messages and prepare data for further processing. The methodology section (X) offers more details 
on precisely which items of data were collected and why.

@@

Users are free to choose any country on their profile. We record each IP address from which a 
user logs in to the site. By examining the distribution of countries (using the MaxMind GeoIP
database\cite{maxmind}) associated with the unique IP addresses (Figure \ref{fig:cip}), we see that the top 
countries for spammers are significantly different, while the top ham countries are 
virtually unchanged. Furthermore, only 30\% of the IP addresses of spam users who 
claimed to be from the United States actually mapped back to a US-based ISP, with a combined 
46\% indicating a Ghana, Nigeria and Senegal ISP (Figure \ref{fig:usaspam}). This contrasts 
with the 73\% that US-based IP addresses comprised for ham users whose profiles stated that 
they were in the United States (Figure \ref{fig:usaham}). During feature extraction, we also 
generated a boolean feature (IP mismatch) indicating whether the IP-detected country and 
the profile country matched.

@@

It is also anecdotally clear from moderator log messages that spammers re-use their profile photos and 
these photos are often of celebrities or simply stock photos. Fingerprinting and perceptual hash techniques 
such as Marr wavelet, discrete cosine transform (DCT), color histograms, etc., which provided by the 
pHash library\cite{phash}, seem to promising in detecting image re-use.
